Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package

### 2020-10-20 050-examples.Rmd
decide on scope of RNA examples, revise introduction

# Replicating motivating data and example applications

@@

```{r previewChapter050, echo = FALSE, eval = FALSE}
bookdown::preview_chapter('050-some-applications.Rmd')
```

@@

## Ultra High-Dimensional simulation-based correlation hypothesis testing under

@@

## Simulation-based joint probability calculations

```{r densityEvaluation, echo=FALSE, eval=FALSE}
## full workflow simulating RNA-seq data

>>>>

# Example applications for our motivating data

@@

```{r previewChapter050, echo = FALSE, eval = FALSE}
bookdown::preview_chapter('050-examples.Rmd')
```

@@

## Simulation-based UHD joint probability calculations

To conduct statistical inference, via maximum likelihood for example, a critical task is to evaluate the joint probability mass (or density) function:

$$
P( \bf{Y} = \bf{y} ), y_i \in \chi_i.
$$

where $\chi_i$ is the sample space for the $i^{th}$ component of the random vector $\bf{Y}$.
Compact representations with convenient computational forms are rare for high-dimensional constructions, especially with hetereogeneous, correlated marginal distributions (or margins of mixed data types).
Given a large number simulated vectors as produced above, estimated probabilities are readily given by counting the proportion of simulated vectors meeting the desired condition.
In our motivating application, one may ask what is the probability that all genes expressed greater than a certain threshold value $\bf{y_0}$.

Then we estimate

$$
\hat{P}( \bf{Y} >= \bf{ y_0 } ) = \sum_{b=1}^B I( \bf{Y^{(b)}} > \bf{ y_0 } ) / B
$$

where $\bf{Y^{(b)}}$ is the $b^{th}$ simulated vector in a total of $B$ simulation replicates and $I(\dot)$ is the indicator function.
For example, we can estimate from our $B=10,000$ simulated vectors the probability that all genes express more than $\bf{y_0} = 1000$.

```{r densityEvaluation, echo=TRUE, eval=TRUE}
d <- ncol(sim_nbinom)
B <- nrow(sim_nbinom)
threshold <- rep( 1000, d)
mean(apply( sim_nbinom,  1,
           function(X, y0=threshold) {
               all( X > y0) }
           ))
```

## UHD Parameteric bootstrap

@@

### 2020-10-20 index.Rmd
decide on scope of RNA examples, revise introduction

Massive high-dimensional data sets are now commonplace in many areas of scientific inquiry.
As new methods are developed for these data, a fundamental challenge lies in designing and conducting simulation studies to assess operating characteristics of proposed methodology, such as false positive rates, statistical power, and robustness --- often in comparison to existing methods.
Another application lies in the discovering the sampling distribution of test statistic under hypothesized null models in complex scenarios.
Such Monte Carlo studies in the high-dimensional setting often become difficult to conduct with existing algorithms and tools, or require expertise in computing on clusters. 
This is particularly true when simulating massive multivariate, non-normal distributions.

>>>>

Massive high-dimensional data sets are now commonplace in many areas of scientific inquiry.
As new methods are developed for these data, a fundamental challenge lies in designing and conducting simulation studies to assess operating characteristics of proposed methodology, such as false positive rates, statistical power, and robustness --- often in comparison to existing methods.
Further efficient simulation empowers computational strategies (such as a parameteric bootstrap --- simulating from a hypothesized null model) to provide inference in analytically challenging settings.
Such Monte Carlo techniques become difficult for high-dimensional data with the current existing algorithms and tools.
This is particularly true when simulating massive multivariate, non-normal distributions, arising naturally in many fields of study in this era of big data.
